---
title: "Assignment 2"
author: "Coco Donovan"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the Necessary Packages

```{r}
library(RMySQL)
library(dplyr)
library(reshape)
library(reshape2)
library(ggplot2)
```

## Establishing the MySQL Connection

```{r}

mysqlconnection = dbConnect(RMySQL::MySQL(),
                            dbname='MovieRatings',
                            host='localhost',
                            port=3306,
                            user='root',
                            password='Navonod17')
```

## Inspecting the available tables in the database

```{r}
dbListTables(mysqlconnection)
```

## Converting movie_ratings databasze into a R dataframe

```{r}
result = dbSendQuery(mysqlconnection, "select * from movie_ratings")

movie_ratings <- fetch(result)

# I added and id here, because I did not want to take an email (for the 
# purposes of anonymity), but I did want a unique identifier

respondent_id <- c(1:nrow(movie_ratings))

movie_ratings$respondent_id <- respondent_id
```

## Reshaping the data frame

```{r}
# I want to keep this here, because it shows that I really did work hard on this
# and that there is always room to grow!

ratings <- movie_ratings[2:ncol(movie_ratings)]

df <- data.frame(matrix(ncol = 3, nrow = ncol(ratings)*nrow(ratings)))

x <- c("Timestamp","movie_names", "movie_ratings")
colnames(df) <- x

movies_names <- c()
individual_ratings <- c()
timestamps <- c()

for (y in 1:ncol(ratings)) {
  for (x in 1: nrow(ratings)) {
    timestamps <- c(timestamps, movie_ratings$Timestamp[x])
    movies_names <- c(movies_names, colnames(ratings)[y])
    individual_ratings <- c(individual_ratings, ratings[x,y])
  }
}

df$Timestamp <- timestamps
df$movie_names <- movies_names
df$movie_ratings <- individual_ratings
```

## Reshaping the data frame (condensed version)

```{r}
melted_movies <- melt(movie_ratings, Movie_Names <- c("Timestamp", "respondent_id"))

colnames(melted_movies) <- c("Timestamp", "respondent_id", "Movie_Name", "Rating")
```

## Percent of total group who watched each movie

```{r}
percent_watched <- melted_movies %>%
  group_by(Movie_Name) %>%
  summarise(percents = (100.0 * sum(Rating != "N/A"))/n_distinct(respondent_id)) %>%
  arrange(percents)

percent_watched

percent_watched_graph <- ggplot(percent_watched, aes(x=percents, y=Movie_Name)) + 
  geom_bar(stat = "identity") +
  labs(title = "Percent of People Surveyed who had seen Each Movie",
       x = "Percent",
       y = "Movie Title")

percent_watched_graph
```

## Average Rating

```{r}
Avg_Ratings <- melted_movies %>%
  group_by(Movie_Name) %>%
  filter(Rating != "N/A") %>%
  summarise(Avgs = sum(as.double(Rating))/sum(Rating != "N/A"))

Avg_Ratings

Avgs_graph <- ggplot(percent_watched, aes(x=Avg_Ratings$Avgs, y=Movie_Name)) + 
  geom_bar(stat = "identity") +
  labs(title = "Average Ratings by Movie Titles",
       x = "Average Rating from 1 to 5",
       y = "Movie Title")

Avgs_graph
```

## Converting movie_info into a R dataframe

```{r}

#for reference "movie_info" is a quick table I put together and added to my 
#MySQL so that I could work with some outside characteristics 

result = dbSendQuery(mysqlconnection, "select * from movie_info")

movie_info <- fetch(result)
```

## Joining movie_ratings and movie_info for analysis

```{r}
Avg_Ratings_and_Info <- merge(x = Avg_Ratings, y = movie_info, by = "Movie_Name",
                                 all.x = TRUE)

# IMDb does ratings out of 10, whereas we did rankings out of 5

Avg_Ratings_and_Info$Scaled_IMDb_Rating <- Avg_Ratings_and_Info$IMDb_Rating/2.0
 
Percent_Watched_and_Info <- merge(x = percent_watched, y = movie_info, by = "Movie_Name",
                                  all.x = TRUE)
```

## Difference from my survey's average ratings compare to IMDb's ratings

```{r}
Avg_Diffs <- Avg_Ratings_and_Info %>%
  summarise(Movie_Name, Difference = Avgs - Scaled_IMDb_Rating) %>%
  arrange(desc(Difference))

Avg_Diffs

# M3GAN and The Whale having the largest difference from the IMDb Rating makes
# sense as they were the least watched
```



```{r}
# Seems like the folks who answered my survey are more critical than the IMDb 
# average

ggplot(Avg_Diffs, aes(x = Difference, y = Movie_Name, fill = Difference)) +
  geom_bar(stat = "identity") +
  labs(title = "The Difference between the Average Ratings for my Survey Compared with IMDb's Ratings",
       y = "Movie Name")
```
