---
title: "Assignment 2"
author: "Coco Donovan"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

## Introduction:

I conducted a survey where I asked people to rate six new movies (released 
within the last three years). I posted the link the survey on Facebook and 
Linkedin, and I shared the link with close friends and family over text. I 
understand that there are biases involved with each way I attempted to solicit
responses, and for the purposes of gathering data in a timely manner I chose to
move forward collecting the movie ratings data in this manner. I explained that 
the data that they provided would be used for an analysis for my Data Management
and Aquisitions class. The movies I chose were The Menu, The Whale, M3GAN, Black
Panther: Wakanda Forever, Everything Everywhere All At Once, and Knives Out. I 
should note that all of the movies were released in 2022 except for Knives Out 
which was released in 2019. I make this note as this time difference may have
impacted how respondents rated Knives Out (the movie may not have been fresh in
their mind).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the Necessary Packages

```{r}
library(RMySQL)
library(dplyr)
library(reshape)
library(reshape2)
library(ggplot2)
```

## Establishing the MySQL Connection

```{r}

mysqlconnection = dbConnect(RMySQL::MySQL(),
                            dbname='MovieRatings',
                            host='127.0.0.1',
                            port=3306,
                            user='root',
                            password='')
```

## Inspecting the available tables in the database

```{r}
dbListTables(mysqlconnection)
```

## Converting movie_ratings databasze into a R dataframe

```{r}
result = dbSendQuery(mysqlconnection, "select * from movie_ratings")

movie_ratings <- fetch(result)

# I added and id here, because I did not want to take an email (for the 
# purposes of anonymity), but I did want a unique identifier

 colnames(movie_ratings) <- c('Timestamp',
         '[M3GAN]',
         '[The Whale]',
         '[The Menu]',
         '[Black Panther: Wakanda Forever]',
         '[Everything Everywhere All At Once]',
         '[Knives Out]')

respondent_id <- c(1:nrow(movie_ratings))

movie_ratings$respondent_id <- respondent_id
```

## Reshaping the data frame

```{r}
# I want to keep this here, because it shows that I really did work hard on this
# and that there is always room to grow!

ratings <- movie_ratings[2:ncol(movie_ratings)]

df <- data.frame(matrix(ncol = 3, nrow = ncol(ratings)*nrow(ratings)))

x <- c("Timestamp","movie_names", "movie_ratings")
colnames(df) <- x

movies_names <- c()
individual_ratings <- c()
timestamps <- c()

for (y in 1:ncol(ratings)) {
  for (x in 1: nrow(ratings)) {
    timestamps <- c(timestamps, movie_ratings$Timestamp[x])
    movies_names <- c(movies_names, colnames(ratings)[y])
    individual_ratings <- c(individual_ratings, ratings[x,y])
  }
}

df$Timestamp <- timestamps
df$movie_names <- movies_names
df$movie_ratings <- individual_ratings
```

## Reshaping the data frame (condensed version)

```{r}
melted_movies <- melt(movie_ratings, Movie_Names <- c("Timestamp", "respondent_id"))

colnames(melted_movies) <- c("Timestamp", "respondent_id", "Movie_Name", "Rating")
```

## Percent of total group who watched each movie

```{r}
percent_watched <- melted_movies %>%
  group_by(Movie_Name) %>%
  summarise(percents = (100.0 * sum(Rating != "N/A"))/n_distinct(respondent_id)) %>%
  arrange(percents)

percent_watched

percent_watched_graph <- ggplot(percent_watched, aes(x=percents, y=Movie_Name)) + 
  geom_bar(stat = "identity") +
  labs(title = "Percent of People Surveyed who had seen Each Movie",
       x = "Percent",
       y = "Movie Title")

percent_watched_graph
```

## Average Rating

```{r}
Avg_Ratings <- melted_movies %>%
  group_by(Movie_Name) %>%
  filter(Rating != "N/A") %>%
  summarise(Avgs = sum(as.double(Rating))/sum(Rating != "N/A"))

Avg_Ratings

Avgs_graph <- ggplot(percent_watched, aes(x=Avg_Ratings$Avgs, y=Movie_Name)) + 
  geom_bar(stat = "identity") +
  labs(title = "Average Ratings by Movie Titles",
       x = "Average Rating from 1 to 5",
       y = "Movie Title")

Avgs_graph
```

## Converting movie_info into a R dataframe

```{r}

#for reference "movie_info" is a quick table I put together and added to my 
#MySQL so that I could work with some outside characteristics 

result = dbSendQuery(mysqlconnection, "select * from movie_info")

movie_info <- fetch(result)
```

## Joining movie_ratings and movie_info for analysis

```{r}
Avg_Ratings_and_Info <- merge(x = Avg_Ratings, y = movie_info, by = "Movie_Name",
                                 all.x = TRUE)

Avg_Ratings_and_Info

# IMDb does ratings out of 10, whereas we did rankings out of 5

Avg_Ratings_and_Info$Scaled_IMDb_Rating <- Avg_Ratings_and_Info$IMDb_Rating/2.0
 
Percent_Watched_and_Info <- merge(x = percent_watched, y = movie_info, by = "Movie_Name",
                                  all.x = TRUE)
```

## Difference from my survey's average ratings compare to IMDb's ratings

```{r}
Avg_Diffs <- Avg_Ratings_and_Info %>%
  summarise(Movie_Name, Difference = Avgs - Scaled_IMDb_Rating) %>%
  arrange(desc(Difference))

Avg_Diffs

# M3GAN and The Whale having the largest difference from the IMDb Rating makes
# sense as they were the least watched
```



```{r}
# Seems like the folks who answered my survey are more critical than the IMDb 
# average

ggplot(Avg_Diffs, aes(x = Difference, y = Movie_Name, fill = Difference)) +
  geom_bar(stat = "identity") +
  labs(title = "The Difference between the Average Ratings for my Survey Compared with IMDb's Ratings",
       y = "Movie Name")
```


## Conclusions:

Everything Everywhere All At Once was by far the most popular movie amongst the 
respondents of my survey. Everything Everwhere All At Once was the most highly
rated movie on average at around 4.2 out of 5, and it was in a three way tie 
as the most viewed movie with around 68% of respondents having seen the film. 
When I compared the ratings from my survey to IDMB ratings, I found that
almost every single movie's average rating from my survey was below the IMDB
rating, expect for average rating for Everything Everywhere All At Once. Part of
the reason the movies may have had low average ratings may have been because there
were so few respondents to my survey compared to the ratings for IMDB. When the 
number of respondents is lower, the mean is more sensitive to change from outliers.
If I wanted to recollect this data, I would ensure a larger data collection window
thereby increasing the chances that I will get a higher amount of responses.